---
title: "GamesAppleStore"
author: "Diego Toralles Avila"
date: "30/10/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


# Introduction

This notebook presents an analysis of the data on 17007 strategy games available on the Apple App Store, such as Clash of CLans, Plants vs Zombies, Pokemon GO and others. This dataset was acquired from Kaggle.com, and it was collected on the 3rd of August 2019 using the iTunes API. 

With this dataset, we may be able to analyze what factors make a sucessful game.


# Loading the dataset

To start this analysis, we first load the required packages (tidyverse, readr) and read the csv file provided by Kaggle.

```{r message=FALSE, warning=FALSE}
if(!require(tidyverse)){install.packages("tidyverse")}
if(!require(readr)){install.packages("readr")}
if(!require(DT)){install.packages("DT")}
options(scipen=10000)

appstoreGamesFile = "data/appstore_games.csv"
appstoreGamesDF = read_csv(appstoreGamesFile) %>% rename_all(~str_replace_all(., "\\s+", ""))
summary(appstoreGamesDF)

```

As seen by the summary, there are 18 columns in this dataset:

- URL: The URL of the game on the Apple App Store.
- ID: The assigned ID of the game on the Apple App Store
- Name: The name of the game.
- Subtitle: The secondary text under the name
- Icon URL:  The URL to the App Icon of the game (a 512px x 512px jpg)
- Average User Rating: Rounded to nearest .5, requires at least 5 ratings
- User Rating Count: Number of ratings internationally, null means it is below 5
- Price: Price in USD
- In-app Purchases: Prices of available in-app purchases
- Description: App description
- Developer: App developer
- Age Rating: Either 4+, 9+, 12+ or 17+
- Languages: ISO2A language codes
- Size: Size of the app in bytes
- Primary Genre: Main genre
- Genres: Genres of the app
- Original Release Date: When it was released
- Current Version Release Date: When it was last updated


We need to fix the typing of some columns, such as the release dates.

```{r}
fixedAppstoreGamesDF <- appstoreGamesDF %>%
  mutate(OriginalReleaseDate = as.Date(OriginalReleaseDate, "%d/%m/%Y")) %>%
  mutate(CurrentVersionReleaseDate = as.Date(CurrentVersionReleaseDate, "%d/%m/%Y")) %>%
  mutate(AgeRating = factor(AgeRating, levels=c('4+','9+', '12+', '17+')))
  
appstoreGamesDF <- fixedAppstoreGamesDF
datatable(appstoreGamesDF %>% select(-URL, -ID, -Subtitle, -IconURL, -Description))
summary(appstoreGamesDF)
```


# Univariate Plots

Right now I have no hypotheses to check, but lets create some plots to see the current state of the games released on the app store. 

## Number of games released each year

First, the number of games released each year. We can see by the plot that the number of games released had been increasing up until 2016. 2017 and 2018 had fewer games released. 2019 is not yet over, so it may catch up to the previous years. 

```{r}
 appstoreGamesDF %>%
  select(OriginalReleaseDate) %>%
  mutate(OriginalReleaseYear = format(OriginalReleaseDate, "%Y")) %>%
  group_by(OriginalReleaseYear) %>%
  summarise(count = n()) %>%
    ggplot(aes(x=OriginalReleaseYear, y=count)) +
    geom_col() +
    geom_text(aes(label=count), vjust=-0.25) + 
    ylab("Number of games released") +
    xlab("Year of Release") +
    theme_minimal()
```

## Number of games with the last update released each year

Ploting the release year of the current version of the games doesn't give us much information. At best, we can see that the majority of the games have had an update in the last 4 years.

```{r}
 appstoreGamesDF %>% select(-URL, -ID, -Subtitle, -IconURL, -Description) %>%
  select(CurrentVersionReleaseDate) %>%
  mutate(CurrentVersionRelease = format(CurrentVersionReleaseDate, "%Y")) %>%
  group_by(CurrentVersionRelease) %>%
  summarise(count = n()) %>%
    ggplot(aes(x=CurrentVersionRelease, y=count)) +
    geom_col() +
    geom_text(aes(label=count), vjust=-0.25) + 
    scale_y_continuous(expand = expand_scale(mult=c(0,0.05))) +
    theme_minimal()
```
## Number of games per user rating.

In this plot, we can see that the number of games per possible rating score increases in a curved fashin up until the 4.5 score. The perfect 5 score is much less common that 4 and 4.5. 

```{r}
unique(appstoreGamesDF$AverageUserRating)

 appstoreGamesDF %>%
  select(AverageUserRating) %>%
  filter(!is.na(AverageUserRating)) %>%
  group_by(AverageUserRating) %>%
  summarise(count = n()) %>%
    ggplot(aes(x=AverageUserRating, y=count)) +
    geom_col() +
    geom_text(aes(label=count), vjust=-0.25) +
    scale_x_continuous(breaks = seq(1,5,by=0.5)) +
    scale_y_continuous(expand = expand_scale(mult=c(0,0.05))) +
    theme_minimal()
```

## Number of games per Age Rating.

This plot is simple, it shows that games on the appstore tend to target all ages. Adult games (17+) are rare.

```{r}
 appstoreGamesDF %>%
  select(AgeRating) %>% 
  arrange(AgeRating) %>%
  group_by(AgeRating) %>%
  summarise(count = n()) %>%
    ggplot(aes(x=AgeRating, y=count)) +
    geom_col() +
    geom_text(aes(label=count), vjust=-0.25) + 
    theme_minimal()
```
## Number of Games per Language.

A game on the appStore may be localized in multiple languages. According to this plot, the two most popular languages for games are English (EN) and Chinese (ZH). This likely reflects the language proficiency of the userbase.

```{r}
 appstoreGamesDF %>%
  select(ID, Languages) %>%
  separate_rows(Languages, sep=",") %>%
  drop_na(Languages) %>%
  group_by(Languages) %>%
  summarise(count = n()) %>%
  arrange(desc(count)) %>%
  top_n(20) %>%
    ggplot(aes(x=reorder(Languages,desc(count)), y=count)) +
    geom_col() +
    geom_text(aes(label=count), vjust=-0.25, size=3.5) +
    scale_y_continuous(expand = expand_scale(mult=c(0,0.05))) +
    theme_minimal()
```

## Number of Games per Genre

Similar to languages, a game mat have multiple genres. The two most common genres are "Strategy" and "Games", which is natural since the dataset we are analyzing is about "Strategy Games". Many games are also classified as "Entertainment", which is not a game genre. The actual most popular game genre in this dataset is "Puzzle".

```{r}
 appstoreGamesDF %>%
  select(ID, Genres) %>%
  separate_rows(Genres, sep=",") %>%
  drop_na(Genres) %>%
  group_by(Genres) %>%
  summarise(count = n()) %>%
  arrange(desc(count)) %>%
  top_n(20) %>%
    ggplot(aes(x=reorder(Genres,desc(count)), y=count)) +
    geom_col() +
    geom_text(aes(label=count), vjust=-0.25, size=3.5) +
    scale_y_continuous(expand = expand_scale(mult=c(0,0.05))) +
    theme_minimal() +
    theme(axis.text.x = element_text(angle=90,vjust= 0.2,hjust=1))
```
## Number of free/not-free games 

The majority of games in the appstore are free to play, as seen in this plot.

```{r}
  appstoreGamesDF %>%
  select(Price) %>%
  drop_na(Price) %>%
  mutate(PriceRange = case_when(Price <= 0 ~ "Free",
                                TRUE ~ "Not Free"))%>%
  mutate(PriceRange = factor(PriceRange, levels= c("Free", "Not Free"))) %>%
  group_by(PriceRange) %>%
  summarise(count = n()) %>%
  ggplot(aes(x=PriceRange, y=count))+
  geom_col()+
  geom_text(aes(label=count), vjust=-0.25, size=3.5) +
  scale_y_continuous(expand = expand_scale(mult=c(0,0.05))) +
  theme_minimal()
```

## Other plots

Finally, I also tried to plot a histogram of the number of games per the number of user ratings However, the histograms is severely unbalanced, that is, most games have very low amounts of user ratings.

```{r}
  appstoreGamesDF %>%
  select(UserRatingCount) %>%
  filter(!is.na(UserRatingCount))%>%
  filter(UserRatingCount>=10000)%>%
  arrange(UserRatingCount) %>%
    ggplot(aes(x=UserRatingCount)) +
    geom_histogram(bins=10) +
    theme_minimal()
```



# Multivariate plots

## What are the most popular genres by age rating?

I removed the "Games", "Entertainment" because they are not game genres. I also removed "Strategy" because the dataset is about games in this genre.

```{r}
 appstoreGamesDF %>%
  select(ID,AgeRating, Genres) %>%
  separate_rows(Genres, sep=",", convert = TRUE) %>%
  mutate(Genres = trimws(Genres)) %>%
  filter(Genres != "Strategy" & Genres != "Games" & Genres != "Entertainment") %>%
  drop_na(Genres) %>%
  group_by(AgeRating,Genres) %>%
  summarise(count = n()) %>%
  arrange(desc(count)) %>%
  top_n(n=5) %>%
  #summarise(averageNumberOfLanguages = mean(numberOfLanguages)) %>%
    ggplot(aes(x=AgeRating, y=count, fill=Genres)) +
    geom_col(position = position_dodge(), width=0.9) +
    #geom_text(aes(label=averageNumberOfLanguages), vjust=-0.25, size=3.5) +
    scale_fill_brewer(palette="Set1") +
    scale_y_continuous(expand = expand_scale(mult=c(0,0.05))) +
    theme_minimal()
```

##   What are the most popular languages by age rating?

```{r}
 appstoreGamesDF %>%
  select(ID,AgeRating, Languages) %>%
  separate_rows(Languages, sep=",", convert = TRUE) %>%
  mutate(Languages = trimws(Languages)) %>%
  # filter(Genres != "Strategy" & Genres != "Games" & Genres != "Entertainment") %>%
  drop_na(Languages) %>%
  group_by(AgeRating,Languages) %>%
  summarise(count = n()) %>%
  arrange(desc(count)) %>%
  top_n(n=7) %>%
    ggplot(aes(x=AgeRating, y=count, fill=Languages)) +
    geom_col(position = position_dodge(), width=0.9) +
    geom_text(aes(label=Languages), vjust=-0.25, size=3.5, position = position_dodge(0.9)) +
    scale_fill_brewer(palette="Set1") +
    scale_y_continuous(expand = expand_scale(mult=c(0,0.05))) +
    theme_minimal() +
    theme(legend.position = "none")
```

We know that English is clearly the most popular language, followed by Chinese (ZH). Since it's not possible to see the difference between the other language columns, lets create the same plot without English and Chinese.

```{r}
 appstoreGamesDF %>%
  select(ID,AgeRating, Languages) %>%
  separate_rows(Languages, sep=",", convert = TRUE) %>%
  mutate(Languages = trimws(Languages)) %>%
  filter(Languages != "EN" & Languages != "ZH") %>%
  drop_na(Languages) %>%
  group_by(AgeRating,Languages) %>%
  summarise(count = n()) %>%
  arrange(desc(count)) %>%
  top_n(n=7) %>%
    ggplot(aes(x=AgeRating, y=count, fill=Languages)) +
    geom_col(position = position_dodge(), width=0.9) +
    geom_text(aes(label=Languages), vjust=-0.25, size=3.5, position = position_dodge(0.9)) +
    scale_fill_brewer(palette="Set1") +
    scale_y_continuous(expand = expand_scale(mult=c(0,0.05))) +
    theme_minimal() +
    theme(legend.position = "none")
```


# Questions

Unfortunately, there is no information regarding the revenue these games make. We can only speculate that any user that reviews a non-free game has bought it at least once. Thus, we can have model of how much money a game has made compared to others. Of course, this does not consider games with in-app purchases, which is not only the the most common type of game in the Apple Store, but they are also the games that usually make the most amount of money in the mobile gaming community according to the news.

With this crude model, we can relate how most variables impact the revenue of a game: e.g., the amount of languages, a specific language, the genres, the release date, the age rating, the app size, and maybe others.

## Is there a correlation between age rating and the number of languages available.

Unfortunatly it is quite difficult to visualize the distribution between these two variables. The values of quartiles are very near one another, so the boxplots are not really useful. The distribution of languages is better shown by ploting a point for each game with some jitter. It was also necessary to crop the y-scale, as some games support more than 90 languages. 

```{r}
  # appstoreGamesDF %>%
  #  select(ID,AgeRating, Languages) %>%
  #  separate_rows(Languages, sep=",") %>%
  #  drop_na(Languages) %>%
  #  group_by(ID,AgeRating) %>%
  #  summarise(numberOfLanguages = n())
 #  arrange(desc(numberOfLanguages))


 appstoreGamesDF %>%
  select(ID,AgeRating, Languages) %>%
  separate_rows(Languages, sep=",") %>%
  drop_na(Languages) %>%
  group_by(ID,AgeRating) %>%
  summarise(numberOfLanguages = n()) %>%
  #ungroup %>%
  #group_by(AgeRating) %>%
  #summarise(averageNumberOfLanguages = mean(numberOfLanguages)) %>%
    ggplot(aes(x=AgeRating, y=numberOfLanguages)) +
    geom_boxplot() +
    geom_jitter(width = 0.3) +
    #geom_text(aes(label=averageNumberOfLanguages), vjust=-0.25, size=3.5) +
    scale_y_continuous(expand = expand_scale(mult=c(0,0.05))) +
    coord_cartesian(ylim=c(0,90)) +
    theme_minimal()
```



## Is there a correlation between age rating and the number of genres?

Similarly to the previous plot, the best way I found to visualize this question is by plotting points for each game. In this case, it seems that age rating has no inflence on the number of genres. The reduced number of points in higher age ratings probably reflects the number of games per each rating.

```{r}
 appstoreGamesDF %>%
  select(ID,AgeRating, Genres) %>%
  separate_rows(Genres, sep=",") %>%
  drop_na(Genres) %>%
  group_by(ID,AgeRating) %>%
  summarise(numberOfGenres = n()) %>%
  #summarise(averageNumberOfLanguages = mean(numberOfLanguages)) %>%
    ggplot(aes(x=AgeRating, y=numberOfGenres)) +
    geom_boxplot() +
    geom_jitter(width = 0.3) +
    #geom_text(aes(label=averageNumberOfLanguages), vjust=-0.25, size=3.5) +
    scale_y_continuous(expand = expand_scale(mult=c(0,0.05))) +
    theme_minimal()
```




### Is there a correlation between age rating and average user rating?

To compare the User ratings for each Age Rating category,  I summed the total amount of user ratings for each rating level and then calculated the ratio of that amount to the total amount of user ratings. This is displayed in the stacked bar chart below.

```{r}
appstoreGamesDF %>%
  drop_na(AverageUserRating) %>%
  arrange(AverageUserRating) %>%
  pull(AverageUserRating) %>%
  unique() -> AverageUserRatingLevels #Get a vector containing all possible user rating levels in sequential order.

appstoreGamesDF %>%
  select(ID,AgeRating, AverageUserRating) %>%
  drop_na(AverageUserRating) %>%
  mutate(AverageUserRating = factor(AverageUserRating, levels = AverageUserRatingLevels)) %>%
  group_by(AgeRating, AverageUserRating) %>%
  summarise(count = n()) %>%
  mutate(freq = count / sum(count)) %>%
  ggplot(aes(x=reorder(AgeRating,desc(AgeRating)), y=freq, fill=AverageUserRating)) + 
  geom_col(position = position_stack(reverse = TRUE))  +
  scale_fill_brewer(palette = "RdYlGn") +
  geom_text(aes(label=count), size=4 ,position=position_stack(vjust = .5, reverse = TRUE)) + 
  theme_minimal() +
  xlab("Age Rating") +
  ylab("Proportion (%)") +
  labs(fill="Average\nUser Rating") +
  coord_flip()
```


# Other questions to explore

There were other questions I thought that could have been analyzed/plotted: 

Is there a correlation between age rating and user rating count?

Is there a correlation between price and age rating?
... between price and the presence of In-app Purchases,
... between price and user rating count?
... between price and language?
... between prices and genre?
... between price and release date?
... between price and AppSize?

Is there a correlation between original release date and current version release date?
... original release date and AppSize.
... original release date and Genre
... original release date and Language







